#Introduction

It is now possible to collect a large amount of data about personal movement using activity monitoring devices such as a Fitbit, Nike Fuelband, or Jawbone Up. These type of devices are part of the “quantified self” movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. But these data remain under-utilized both because the raw data are hard to obtain and there is a lack of statistical methods and software for processing and interpreting the data.

This assignment makes use of data from a personal activity monitoring device. This device collects data at 5 minute intervals through out the day. The data consists of two months of data from an anonymous individual collected during the months of October and November, 2012 and include the number of steps taken in 5 minute intervals each day.

```{r}

activity<-read.csv("/Users/eshanchatty/Downloads/activity.csv")
sum(is.na(activity))
library(ggplot2)
activity <- data.table::fread(input = "/Users/eshanchatty/Downloads/activity.csv")
sum(is.na(activity))
activity<-na.omit(activity)
```
Calculate the total steps taken per day.
```{r}
Total_Steps <- activity[, c(lapply(.SD, sum, na.rm = FALSE)), .SDcols = c("steps"), by = .(date)] 
head(Total_Steps)
```
Plot a histogram of number of steps taken each day.
```{r}
ggplot(Total_Steps, aes(x = steps)) +
    geom_histogram(fill = "blue", binwidth = 1000) +
    labs(title = "Daily Steps", x = "Steps", y = "Frequency")
```



Calculate the mean and median of number of steps taken each day.
```{r}
mean(Total_Steps$steps,na.rm = TRUE)
median(Total_Steps$steps,na.rm = TRUE)
```
What is the Average daily pattern?
```{r}
Avgpattern <- activity[, c(lapply(.SD, mean, na.rm = TRUE)), .SDcols = c("steps"), by = .(interval)] 

ggplot(Avgpattern, aes(x = interval , y = steps)) + geom_line(color="blue", size=1) + labs(title = "Avg. Daily Steps", x = "Interval", y = "Avg. Steps per day")
```




1 -> Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.
#Replacing the NA values with the median, as follows.
```{r}
activity[is.na(steps), "steps"] <- activity[, c(lapply(.SD, median, na.rm = TRUE)), .SDcols = c("steps")]
```
2 -> Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
data.table::fwrite(x = activity, file = "/Users/eshanchatty/Downloads/tidyData.csv", quote = FALSE)
```
3 -> Make a histogram of the total number of steps taken each day and calculate and report the mean and median total number of steps taken per day. 

```{r}
Total_Steps <- activity[, c(lapply(.SD, sum)), .SDcols = c("steps"), by = .(date)] 

Total_Steps[, .(Mean_Steps = mean(steps), Median_Steps = median(steps))]
```
Plot:
```{r}
ggplot(Total_Steps, aes(x = steps)) + geom_histogram(fill = "blue", binwidth = 1000) + labs(title = "Daily Steps", x = "Steps", y = "Frequency")
```




4 -> Do these values differ from the estimates from the first part of the assignment? <br>
Yes, they do differ.<br>
5 -> What is the impact of imputing missing data on the estimates of the total daily number of steps?<br>
The Data with NA values has the mean=median=10765.<br>
The Data without NA values has mean = 9354.23 and Median = 10395
Hence, the mean and median has shifted to the left i.e it is now right skewed.<br>
6 -> Are there differences in activity patterns between weekdays and weekends?
```{r}
activity[, date := as.POSIXct(date, format = "%Y-%m-%d")]
activity[, `Day of Week`:= weekdays(x = date)]
activity[grepl(pattern = "Monday|Tuesday|Wednesday|Thursday|Friday", x = `Day of Week`), "weekday or weekend"] <- "weekday"
activity[grepl(pattern = "Saturday|Sunday", x = `Day of Week`), "weekday or weekend"] <- "weekend"
activity[,`weekday or weekend` := as.factor(`weekday or weekend`)]
activity[is.na(steps), "steps"] <- activity[, c(lapply(.SD, median, na.rm = TRUE)), .SDcols = c("steps")]
Interval <- activity[, c(lapply(.SD, mean, na.rm = TRUE)), .SDcols = c("steps"), by = .(interval, `weekday or weekend`)] 

ggplot(Interval , aes(x = interval , y = steps, color=`weekday or weekend`)) + geom_line() + labs(title = "Avg. Daily Steps by Weektype", x = "Interval", y = "No. of Steps") + facet_wrap(~`weekday or weekend` , ncol = 1, nrow=2)
```
Hence Clearly we can see that on weekends the number of steps are clearly way more higher and the interval too is high.
